
# My fried received a simple report on the status of civic society in the UK before the London Bombings. The report consists of numbers that my friend does not understand and she needs to report to the high command. First I would like to help her review the data by loadig it.
```{r}
load(url("http://jakebowers.org/Data/ho05.rda")) # import a dataset
# Since she is interested in the data before the London Bombing, I would look at the column of "postbomb". As given from the codebook, variables that have the code "0" fall under our observation. 
table(ho05$postbomb,useNA="ifany") # creating a frequency table for the variable postbomb, useNA="ifany" includes in the table the counts of NA values if there is any missing values  
# Since 800 variables are mixed with other 200 variables, I would like to create another data sheet that consists of variables with code "0". 
library(dplyr) # loading package that provides data manipulating functions 
wrkdat <- ho05  %>% filter(postbomb==0) # creating data wrkdat, filter results to include the value of postbomb that equals to 0
# Looking at the codebook, I could understand what the column "hlphrs" signifies. Apparently, the program interviewd people on how many hours they spent helping some groups, clubs, or organizations and how many hours they spent helping other people as an individual. The variable "hlphrs" is the sum of group hours and individual hours.
sum(wrkdat$hlphrs, na.rm=T) # sum all the values of the variable hlphrs, which gives the result of NA due to  missing values. In order to count the sum, we can add the function na.rm=T.  
```


\end{verbatim}
\end{small}

#Referring to the question that my friend asked, "What is the best way to say how civic life was going before the bombings? What is the right answer?", I would suggest that the variable "hlphrs" is very informative in terms of how people engaged in civic life. However, presenting all the values of the column is not the best way. I would not suggest describing the variable in one dimension, which is giving one value that represents the whole data.
## The best way to describe the civic life by using the given variable, help hours, would be to find the proper representation. There are several ways to describe the variable, for example, minimum and maximum, median, and the mean. Obviously, there are various ways to calculate the mean. There is no right answer, but it differs according to how we want to treat the outliers.


# The codes help us to know the average time that people spend to help others. By controlling outliers we can obtain different means to describe a single variable. Depending on the purpose of research, the importance of outliers and how we control them varies. By calculating different kinds of means we can better analyze and interpret data. 

# histogram could help to visualize the given data, code: hist(wrkdat$hlphrs)

```{r, results='hide'}
# When we cacaluate the mean it is 6.633626, which might not be representing the data as it could be influenced by outliers. Therefore, we use other methods such as the trimmed mean and winsorized mean.
mean(wrkdat$hlphrs,trim=.1,na.rm=TRUE) #mean of the variable hlphrs after removing the lowest and highest 5% values, na.rm=T omits all NA from the calculation. There is a difference between the trimmed mean (3.327074) and the mean (6.633626). 
hlp <- unlist(wrkdat %>% dplyr::select(hlphrs) %>% filter(!is.na(hlphrs)))
# create values of hlp
# from wrkdat data select hlphrs and show all non-missing values in the data
# The following is the procedures of obtaining two different kinds of means that adjust the affect of outliers.
library(psych) #loading package psych
winsor.mean(hlp) #calulating mean after replacing extreme values. The result is 3.134253.
library(robustHD) # loading package robustHD
mean(winsorize(hlp)) # winsorize (replacing the extreme values) of hlp and calulating the mean. The result is 1.802459

# The following is the functions to obtain the one step Huber M estimator. Rather than symmetrically trimming the predetermined amount of the data, we can adjust the amount of trimming based on the distribution of outliers by using this method.
onestepMest <- function(x){
	## Following http://www.psychology.mcmaster.ca/bennett/boot09/rt2.pdf
	madn <- mad(x, constant = 1.4826) ## 1/1.4826 = .6745   
  #mad calculates the median absolute deviations from the median of a dataset. It is a robust measure of the variability in a sample
	#The result is stored as madn. 	
	M <-  ( abs(x - median(x)) / madn )  > 1.28 
	#abs computes the absolute value of (x-median(x)). The result is divided by madn. 
  U <- sum( x > M) #summing values that are greater than M. The result is named U
	L <- sum( x < M) # summing values that are less than M.The result is named L
	B <- length(x) - U - L  
	n <- length(x) 
	mux <- (1.28 * madn * (U-L) + B) /(n - L - U) 
	return(mux) # return the value of mux 
}

onestepMest(hlp) #insert hlp to function named 'onestepMest' 
library(robustbase) #loading package
huberM(hlp) #driving huberM estimator  
fivenum(hlp) #returns the summary of five functions (minimum, lower-hinge, median, upper-hinge, maximum)
quantile(hlp,seq(0,1,.1))#quantile of hlp, with a sequence of 10% interval 

```

# There are various ways to gain information of a data by calcuating different means. There are different methods to control outliers that could distort the interpretation of data. The trimmed mean and mean obtained by the function onestepMest compute similar estimates of mean. By comparing the different outcomes we can better understand the data.

<!-- see also https://dornsife.usc.edu/labs/rwilcox/software/ and WRS:::mest  https://dornsife.usc.edu/assets/sites/239/docs/WRS2.pdf and the MASS library--->

# References
